Robust inference of neutron-star parameters from thermonuclear burst

### 2021-09-14 doc/ms.tex
Added homogenize_params to help ensure compatible parameter types
homogenize_params checks the size of all the passed parameters, and if any of them are a distribution, will convert the others to distributions of the same size (by calling value_to_dist). So far the alpha and mdot methods have been converted to use this routine, and the others should also be converted.

Also added logging (in utils.py) via the logging package, to make the informational and error messages look nicer

Thermonuclear (type-I) bursts arise from unstable ignition of accumulated fuel on the surface of neutron stars in low-mass X-ray binaries. Measurements of burst properties enable observers to infer the properties of the host neutron stars, but a number of confounding astrophysical effects contribute to systematic uncertainties. 
%
In particular, comparison of observed burst properties with the predictions of numerical models is a 
% key factor in determining the properties of the host neutron star.
powerful method of constraining host properties.
%
Here we describe some commonly-used approaches for determining system parameters, including composition of the burst fuel, and introduce a new suite of software tools, {\sc concord}, intended to make comprehensive model-observation comparisons practical.

\end{abstract}

@@

Here we explore the ways that host system  parameters affect the observed properties of thermonuclear bursts, and assess to what degree these effects can be corrected for.
%
In \S\ref{sec:data} we describe the typical characteristics of burst observations along with those properties that can be 
% predicted by numerical models.
inferred from simple calculations or by comparison with numerical models.
% 
In \S\ref{sec:analysis} we explore a number of methods to infer properties of the bursting source, based on the extent of available observational data, and assess their accuracy.
%
In \S\ref{sec:simulations} we apply these methods to a range of simulated data to assess what degree of systematic errors these measurements might be subject to.

@@

% ------------------------------------------------------------------------------
\section{Observations and modelling} 
\label{sec:data}

%\subsection{Burst observations}
%\label{subsec:data}

The fundamental observable that we measure is the time-history of the burst flux, the lightcurve $(t_i,F_{i})$. Here the $F_{i}$ may be in instrumental units of count~s$^{-1}$ (or perhaps count~s$^{-1}$~cm$^{-2}$), or may be estimates of the bolometric flux at time $t_i$, derived from model fits of the spectrum at time $t_i$.
%
Provided the $F_i$ are in units of flux, or the flux can be estimated from the intensity,  we may measure the peak flux, $F_{\rm pk}$, and integrate to estimate the burst fluence, $E_b$.

Given one or more bursts, we can derive constraints on the recurrence time $\Delta t$. A pair of bursts separated by an interval with complete coverage by X-ray instruments offer an unambiguous measurement of the recurrence time, but such measurements are rare due to the typically low duty cycle for most instruments \cite[e.g.][]{minbar}.
%
Such cases also tend to be biased towards lower recurrence times, which may be dominated by so-called ``short waiting time'' bursts, which likely arise from ignition of unburnt fuel from the previous event \cite[]{keek10}. The recurrence time for such events likely does not represent the actual time required to meet the ignition conditions at the base of the fuel layer, and so are generally excluded for the purposes of inferring properties of the burst sources.

Where one or more bursts $N$ are observed in low-duty cycle observations, the rate can be estimated based on the total exposure, i.e.
\begin{equation}
\langle R\rangle = \frac{N}{\sum^n T_i}
\end{equation}
where each of $n$ observartions has exposure $T_i$. The probable range of the average rate can be estimated assuming Poisson counting statistics for the uncertainty on the burst number $N$. Because of the approximately periodic behavior common for bursts, and the similarity between low-Earth orbital periods and the burst recurrence times, such uncertainty estimates may be wildly incorrect.

From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; 
% and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
%
these parameters may be combined to estimate the $\alpha$-value, the (observed) ratio of the burst to persistent flux: 
\begin{equation}
\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
\end{equation}
where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.

@@

A significant fraction of observed bursts exhibit characteristic variations in their spectral hardness (or the blackbody temperature and normalisation) around their maximum that indicate the presence of photospheric radius-expansion (PRE). 
%
Even if such variations are not present, 
%
the peak flux $F_{\rm pk}$ provides constraints on the source distance, since the maximum burst luminosity is limited to (roughly) the Eddington luminosity \cite[e.g.][]{lew93}:
% copied from bcatalog, and updated prefac
% (8.*pi*const.G*const.m_p*1.4*const.M_sun*const.c/(const.sigma_T*1.259)).to('erg s-1')                                                         
% Out[16]: <Quantity 2.79569695e+38 erg / s>
\begin{eqnarray}
  L_{\rm Edd,\infty} & = & \frac{8\pi G m_p M_{\rm NS} c
  [1+(\alpha_{\rm T}T_{\rm e})^{0.86}]} {\sigma_T(1+X)[1+z(R)]} % \xi_b
       \nonumber \\
%  & = & 2.7\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
  & = & 2.80\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
 \frac{1+(\alpha_{\rm T}T_{\rm e})^{0.86}}{(1+X)}
\nonumber \\ & & \times\  % line break
%    \left[\frac{1+z(R)}{1.31}\right]^{-1}\ % \xi_b\
    \left[\frac{1+z(R)}{\opz}\right]^{-1}\ % \xi_b\
              \eps
  \label{ledd}
\end{eqnarray}
Here
$T_{\rm e}$ is the
effective temperature of the atmosphere, $\alpha_{\rm T}$ parametrizes the temperature dependence of the electron scattering
opacity \cite[$\simeq 2.2\times10^{-9}$~K$^{-1}$;][]{lew93},
$m_p$ is the proton mass, $\sigma_T$ the Thompson 
cross-section, and $X$ the
mass fraction of hydrogen in the atmosphere.
% ($\approx0.7$ for cosmic abundances).
The final factor in
square brackets represents the gravitational redshift 
at the photosphere $1+z(R)=(1-2GM_{\rm NS}/R
c^2)^{-1/2}=1.259$ for $M_{\rm NS}=1.4\ M_\odot$ and $R=R_{\rm NS}=11.2$~km.
%
The effective redshift measured at the peak of a PRE burst may be lower than the value at the NS surface, while the photosphere is expanded during the radius expansion episode (i.e. $R\ge R_{\rm NS}$).

@@

\subsection{Emission anisotropy}
\label{subsec:anisotropy}

@@

@@

@@

In order to test the analysis approaches described in this paper, we generated sets of simulated data based on the {\sc kepler} models described in \S\ref{subsec:models}. 
%
This procedure is identical to that used to convert model predictions to data for direct comparison to observations, as described in \S\ref{subsec:lccompare}.
%
We seek to calculate the burst lightcurve that would be observed given a model lightcurve and neutron star parameters including distance $d$; inclination $i$; and surface redshift $(1+z)_{\rm fit}$ (so labeled to distinguish from the corresponding value implict to the model). The process can be summarised as follows. [{\bf note doesn't include any steps to deal with the redshift inconsistency, see below}]
\begin{enumerate}
\item Multiply the time bins for the predicted burst lightcurve by the trial gravitational redshift $(1+z)_{\rm fit}$, thereby ``stretching'' the profile to account for the general relativistic time dilation at the neutron-star surface
% \item Shift the predicted burst lightcurve by $t_{\rm off}$
\item Apply the same correction to the model-predicted recurrence time $\Delta t_{\rm pred}$.
\item Interpolate the model-predicted lightcurve onto a set of observational time bins, corresponding (for example) to the typical resolution for time-resolved spectroscopy (0.25~s)
\item Translate the model-predicted luminosity to the corresponding quantity measured by a distant observer, by 
% multiplying by $[\xi/(1+z)]^2$
dividing by $(1+z)_{\rm fit}$
\item Convert the luminosity to (isotropic) flux by dividing by the distance factor, $4\pi d^2$ \label{distfac}
\item Take into account the expected anisotropy effects due to the system inclination, by dividing the luminosity by the anisotropy factor $\xi_b$ \label{anisofac}
\item Calculate the persistent flux expected for the model-assumed accretion rate $\dot{m}$, and apply the same corrections as for the burst flux in steps \ref{distfac} and \ref{anisofac} (adopting a separate anisotropy factor $\xi_p$ appropriate for the persistent flux). We also divide by a bolometric correction factor $c_{\rm bol}$ accounting for the limited instrumental passband.
\end{enumerate}
% Below we describe each of these steps in more detail

@@

\section{Analysis} 
\label{sec:analysis}

Here we describe different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.
%
We focus on adopting a suitable approach given the availability of data, which will differ for different sources

@@

While it may seem strange to base an analysis on the {\it non-}detection of bursts, practically even for the best-studied sources the X-ray observation duty cycle is of order a few percent. There is a high probability that bursts will be missed, particularly if the accretion rate is low (and hence the bursts are infrequent). In extreme cases this can mean that no bursts whatsoever are observed, which may instead be explained by the compact object being a black hole rather than a neutron star.

@@

For example, based on {\it RXTE}/PCA observations of the accretion-powered millisecond pulsar IGR~J00291+5934, \cite{gal06b} derived joint constraints over the distance $d$ and fuel H-fraction $X_0$, adopting  the predictions of a simple numerical model ({\it ref to {\sc settle}}). These constraints suggest the distance is $\gtrsim5$~kpc (at 3-sigma significance), provided $X_0\approx0.7$, a reasonable choice given the expectation of a H-rich donor in this $2.46$~hr binary orbital period system \cite[]{gal05a}.  \todo{give more detail about how this was done}

In a subsequent outburst in 2015, a single burst was detected by {\it Swift}/XRT \cite[]{kuin15}, offering the opportunity to verify the previously-determined distance limit. The estimated peak 
% 0.3--10~keV flux in the range (3.0--$5.6)\times10^{-11}\ \epcs$. 
0.1--35~keV flux was $(18\pm4)\times10^{-8}\ \epcs$ \cite[]{defalco17}, with the burst exhibiting spectral variations indicative of PRE. By comparing the peak flux to the empirical Eddington luminosity (see \S\ref{subsec:1burst}, below), the corresponding distance 
\cite[including the inclination range of 22--32$^\circ$ suggested by][]{torres08},
% possible inclinations as described in sec \S\ref{subsec:data}) is $4.4_{-0.7}^{+0.8}$~kpc, which is mostly inconsistent with the earlier limit. Clearly, such limits must be viewed with some caution.
is $5.0_{-0.5}^{+0.7}$~kpc, which is fully consistent with the previously established limit. The effect of the inclination constraints is illustrated in Fig. \ref{fig:dist_igr00291}).

\begin{figure}[ht!]
\plotone{fig1.pdf}
\caption{Estimated distance distributions for the accretion-powered X-ray pulsar IGR~J00291+5934, based on the peak flux of the sole thermonuclear burst observed by \cite{defalco17}.
%
The distribution for an isotropic distribution of system inclinations (up to a maximum value of $75^\circ$, motivated by the lack of dips in the X-ray intensity) is shown as the blue shaded histogram. Imposing the inclination constraint of 22--32$^\circ$ suggested by \cite{torres08}, for which the predicted burst anisotropy factor is $\xi_b=0.704$ on average, gives instead the orange-shaded histogram. The resulting $1\sigma$ confidence interval is $5.0_{-0.5}^{+0.7}$~kpc.
\label{fig:dist_igr00291}}
\end{figure}

\subsection{One burst -- estimating the burst rate} 
\label{subsec:1burst}

@@

For a single burst detected from a source, we can expect to measure the burst peak flux, fluence, and timescale, and infer the distance and hence the expected burst rate. The detailed shape of the lightcurve can provide constraints from comparisons with numerical models (see \S\ref{subsec:lccompare}), but here we focus on the more straightforward calculation which may be done from the simple burst measurements alone.

We use for example the sole burst detected with \igr/JEM-X from IGR~J17591$-$2342 \cite[]{kuiper20}. This burst was observed  when the estimated persistent bolometric flux was $(1.2\pm0.2)\times10^{-9}\ \epcs$, and exhibited a fluence of $E_b =(1.1\pm0.1)\times10^{-6}\ {\rm erg\,cm^{-2}}$. The peak flux was $(7.6\pm1.4)\times10^{-6}\ \epcs$.

\cite{kuiper20} first estimated the distance to the source based on the empirical value of the Eddington luminosity, and incorporating the expected anisotropy for an inclination range of 
% kuiper20 uses slightly more restrictive 28-30
24--30$^\circ$ \cite[]{sanna18} as $7.7_{-0.6}^{+0.8}$~kpc. Here the uncertainty is dominated by the peak flux, since the anisotropy factor is effectively fixed for such a narrow range of inclination, at $\xi_b=0.70$.

@@

We can use this distance constraint to estimate in turn the accretion and burst rate, as follows.
%
The persistent flux can then be converted to an accretion rate, given the estimated anisotropy factor predicted (separately) for the persistent emission, as
\begin{eqnarray}
\dot{m} & = L_{\rm pers} (1+z) (4\pi R^2(GM/R_{\rm NS}))^{-1}\\
& = \ldots
% TODO: give expanded expression for this and move to earlier section
\end{eqnarray}
via the {\tt mdot} method, and incorporating the estimated distance distribution already determined. The estimated value is within the range $(2.0_{-0.4}^{+0.6})\times10^3\ \rm{g\,cm^{-2}\,s^{-1}}$ with the usual assumed range for neutron star mass and radius (and hence redshift).

In parallel, we can estimate the ignition column using equation \ref{eq:column}, given an estimate of the nuclear burning yield, $\qnuc$. ; as before, we could potentially obtain an estimate of this quantity by comparison of the observed lightcurve to simulations. However, in the absence of an estimate of this quantity, all is not lost.

Instead we can work backwards, 
% 
and assume a set of randomly generated values of \Xb\
to estimate $\yign$ from equation \ref{column}, and $\Delta t_{\rm rec} = \yign/\dot{m}$.
%
We choose a uniform distribution for \Xb\ in the range 0--0.7; this distribution is probably not realistic for a wide sample of bursts, but it is difficult to do better.
% text copied from Kuiper et al. 2020 down to ********
% not yet edited (need to modify sufficiently to avoid plagiarism issues
Given each pair of (\Xb,  $\Delta t_{\rm rec}$) we can infer the accreted H-fraction $X_0$, based on 
% the expectation that hydrogen burns via the hot-CNO cycle and will be exhausted in a time $t_{\rm CNO} =9.8(X_0/0.7)(Z_{\rm CNO}/0.02)^{-1}$ \citep[e.g.][]{lampe16}.
equation \ref{eq:X_0}.
%
Since $Z_{\rm CNO}$ is also unknown, we similarly draw random values from a uniform distribution in the range 0.0--0.02.
We then select only those $X_0$ values within a physically realistic range $X_0<0.75$, and compute the confidence limits on the remaining parameters.

For the single burst observed from IGR~J17591$-$2342 the inferred ignition column is $2.2_{-0.8}^{+1.4}\times10^8\,{\rm g\,cm^{-2}}$, and the average expected recurrence time is $1.8\pm0.7$~d. 
% Such long recurrence times likely guarantee that the hydrogen in the accreted fuel is exhausted at the base prior to ignition, and explaining the fast rise and short duration of the burst. 
We can also infer lower (upper) limits on $X_0$ ($Z_{\rm CNO}$), although these limits are not strongly constraining; we find $X_0>0.17$ and $Z_{\rm  CNO}<0.017$ at 95\% confidence.
% *******

@@

>>>>

Thermonuclear (type-I) bursts arise from unstable ignition of accumulated fuel on the surface of neutron stars in low-mass X-ray binaries. Measurements of burst properties enable observers to infer the properties of the host neutron star and mass donors, but a number of confounding astrophysical effects contribute to systematic uncertainties. Here we describe some commonly-used approaches for determining system parameters, including composition of the burst fuel, and introduce a new suite of software tools, {\sc concord}, intended to fully account for astrophysical uncertainties. Furthermore, comparison of observed burst properties with the predictions of numerical models is a powerful method of constraining host properties, and the tools presented here are intended to make comprehensive model-observation comparisons straightforward. When combined with the extensive samples of burst observations accumulated by X-ray observatories, these software tools will provide a step-change in the amount of information that can be inferred about typical burst sources.
\end{abstract}

@@

Here we explore the ways that host system  parameters affect the observed properties of thermonuclear bursts, and assess to what degree these effects can be corrected for.
%
In \S\ref{sec:data} we describe the typical characteristics of burst observations along with those properties that can be 
% predicted by numerical models.
inferred from simple calculations. % or by comparison with numerical models.
% 
In \S\ref{sec:analysis} we explore a number of methods to infer properties of the bursting source, based on the extent of available observational data, and assess their accuracy.
%
In \S\ref{sec:simulations} we apply these methods to a range of simulated data to assess what degree of systematic errors these measurements might be subject to.

@@

% ------------------------------------------------------------------------------
\section{Observations and burster properties} 
\label{sec:data}

% describe the typical characteristics of burst observations along with those 
% properties that can be inferred from simple calculations

%\subsection{Burst observations}
\label{subsec:data}

The fundamental observable that we measure is the time-history of the burst flux, the lightcurve $(t_i,F_{i})$. Here the $F_{i}$ may be in instrumental units of count~s$^{-1}$ (or perhaps count~s$^{-1}$~cm$^{-2}$), or may be estimates of the bolometric flux at time $t_i$, derived from spectral model fits of the spectrum at time $t_i$.
%
The net (excluding the persistent emission) burst spectra generally can be well fit with a blackbody model, and it is thought such models can provide a reasonable estimate of the bolometric flux [{\bf ref}]

[{\bf insert table with key burst parameters here? }]

Provided the $F_i$ are in units of flux, or the flux can be estimated from the intensity,  we may measure the peak flux, $F_{\rm pk}$, and integrate to estimate the burst fluence, $E_b$. 
%
Where the bursts are very short, or observed with low signal-to-noise, a spectrum extracted over an interval covering the entire event may substitute for a fluence measurement.

Given one or more bursts, we can derive constraints on the recurrence time $\Delta t$. A pair of bursts separated by an interval with uninterrupted coverage by X-ray instruments offer an unambiguous measurement of the recurrence time, but such measurements are rare due to the typically low duty cycle for most instruments \cite[e.g.][]{minbar}.
%
Such cases also tend to be biased towards lower recurrence times, which may be dominated by so-called ``short waiting time'' bursts, which likely arise from ignition of unburnt fuel from the previous event \cite[]{keek10}. The recurrence time for such events likely does not represent the actual time required to meet the ignition conditions at the base of the fuel layer, and so are generally excluded for the purposes of inferring properties of the burst sources.

Where one or more bursts $N$ are observed in $n$ low-duty cycle observations with exposure $T_i$, the rate can be estimated based on the total exposure, i.e.
\begin{equation}
\langle R\rangle = \frac{N}{\sum^n T_i}
\end{equation}
% where each of $n$ observartions has exposure $T_i$. 
The probable range of the average rate can be estimated assuming Poisson counting statistics for the uncertainty on the burst number $N$. Because of the approximately periodic behavior common for bursts, and the similarity between low-Earth orbital periods and the burst recurrence times, such uncertainty estimates may be wildly incorrect.

One of the most significant astrophysical uncertainties that has largely been neglected to date is the anisotropy of the burst (and persistent) emission.
% [{\bf move discussion from later section to here?}]
% \subsection{Emission anisotropy}
% \label{subsec:anisotropy}
[{\bf bit more description here }]

@@

A significant fraction of observed bursts exhibit characteristic variations in their spectral hardness (or the blackbody temperature and normalisation) around their maximum that indicate the presence of photospheric radius-expansion (PRE) [{\bf ref}]. 
%
Even if such variations are not present, 
%
the peak flux $F_{\rm pk}$ provides constraints on the source distance, since the maximum burst luminosity is limited to (roughly) the Eddington luminosity \cite[e.g.][]{lew93}:
% copied from bcatalog, and updated prefac
% (8.*pi*const.G*const.m_p*1.4*const.M_sun*const.c/(const.sigma_T*1.259)).to('erg s-1')                                                         
% Out[16]: <Quantity 2.79569695e+38 erg / s>
\begin{eqnarray}
  L_{\rm Edd,\infty} & = & \frac{8\pi G m_p M_{\rm NS} c
  [1+(\alpha_{\rm T}T_{\rm e})^{0.86}]} {\sigma_T(1+X)[1+z(R)]} % \xi_b
       \nonumber \\
%  & = & 2.7\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
  & = & 2.80\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
 \frac{1+(\alpha_{\rm T}T_{\rm e})^{0.86}}{(1+X)}
\nonumber \\ & & \times\  % line break
%    \left[\frac{1+z(R)}{1.31}\right]^{-1}\ % \xi_b\
    \left[\frac{1+z(R)}{\opz}\right]^{-1}\ % \xi_b\
              \eps
  \label{ledd}
\end{eqnarray}
Here
$T_{\rm e}$ is the
effective temperature of the atmosphere, $\alpha_{\rm T}$ parametrizes the temperature dependence of the electron scattering
opacity \cite[$\simeq 2.2\times10^{-9}$~K$^{-1}$;][]{lew93},
$m_p$ is the proton mass, $\sigma_T$ the Thompson 
cross-section, and $X$ the
mass fraction of hydrogen in the atmosphere.
% ($\approx0.7$ for cosmic abundances).
The final factor in
square brackets represents the gravitational redshift 
at the photosphere $1+z(R)=(1-2GM_{\rm NS}/R
c^2)^{-1/2}=1.259$ for $M_{\rm NS}=1.4\ M_\odot$ and $R=R_{\rm NS}=11.2$~km.
%
The effective redshift measured at the peak of a PRE burst may be lower than the value at the NS surface, while the photosphere is expanded during the radius expansion episode (i.e. $R\ge R_{\rm NS}$).

@@

@@

% \subsection{Burst energetics}
\label{subsec:energy}

@@

From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; [{\bf add expression here}]
% and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
the persistent spectra of burst sources are  more complex than the bursts themselves, and typically exhibit thermal as well as non-thermal components [{\bf ref}].
%
These parameters may be combined to estimate the $\alpha$-value, the (observed) ratio of the burst to persistent flux: 
\begin{equation}
\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
\end{equation}
where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.
%
The $\alpha$ parameter is an important diagnostic of the burst fuel and ignition conditions.
%
[{\bf mention the average method, as used in the rates paper }]

Given an estimate of $Q_{\rm nuc}$, the average hydrogen fraction $\langle X\rangle$ in the fuel layer at ignition has been estimated to date using the relation $Q_{\rm nuc}=1.6 + 4\langle X\rangle$~MeV/nucleon \cite[e.g.][and references therein]{gal03d}. This expression includes $\approx35$\% losses
attributed to neutrino emission \cite[]{fuji87}. However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain, and recent {\sc kepler} simulations indicate instead an approximation of 
% $Q{\rm nuc}=0.96+6.6\langle X\rangle$ 
$Q{\rm nuc}=1.35+6.05\langle X\rangle$ 
% (Goodwin et al. 2018, in preparation). 
\cite[]{goodwin19a}.
[{\it comment on improved quadratic expression}]
%
The average H-fraction at ignition $\langle X\rangle$ is a lower limit for the H-fraction in the accreted fuel, $X_0$, as for most sources the H-fraction will be reduced by steady burning prior to burst ignition.
%
The recurrence time for the burst can further be used to estimate $X_0$, although the burning rate also depends on the CNO mass fraction, $Z_{\rm CNO}$.

@@

In order to test the analysis approaches described in this paper, we generated sets of simulated data based on the {\sc kepler} models described in \S\ref{subsec:models}. 
%
This procedure is identical to that used to convert model predictions to data for direct comparison to observations, as described in \S\ref{subsec:lccompare}.
%
\todo{if I just use Adelle's data I don't need to mention the lightcurves at all}

We seek to calculate the burst lightcurve that would be observed given a model lightcurve and neutron star parameters including distance $d$; inclination $i$; and surface redshift $(1+z)_{\rm fit}$ (so labeled to distinguish from the corresponding value implict to the model). The process can be summarised as follows. [{\bf note doesn't include any steps to deal with the redshift inconsistency, see below}]
\begin{enumerate}
\item Multiply the time bins for the predicted burst lightcurve by the trial gravitational redshift $(1+z)_{\rm fit}$, thereby ``stretching'' the profile to account for the general relativistic time dilation at the neutron-star surface
% \item Shift the predicted burst lightcurve by $t_{\rm off}$
\item Apply the same correction to the model-predicted recurrence time $\Delta t_{\rm pred}$.
\item Interpolate the model-predicted lightcurve onto a set of observational time bins, corresponding (for example) to the typical resolution for time-resolved spectroscopy (0.25~s)
\item Translate the model-predicted luminosity to the corresponding quantity measured by a distant observer, by 
% multiplying by $[\xi/(1+z)]^2$
dividing by $(1+z)_{\rm fit}$
\item Convert the luminosity to (isotropic) flux by dividing by the distance factor, $4\pi d^2$ \label{distfac}
\item Take into account the expected anisotropy effects due to the system inclination, by dividing the luminosity by the anisotropy factor $\xi_b$ \label{anisofac}
\item Calculate the persistent flux expected for the model-assumed accretion rate $\dot{m}$, and apply the same corrections as for the burst flux in steps \ref{distfac} and \ref{anisofac} (adopting a separate anisotropy factor $\xi_p$ appropriate for the persistent flux). We also divide by a bolometric correction factor $c_{\rm bol}$ accounting for the limited instrumental passband.
\end{enumerate}
% Below we describe each of these steps in more detail

@@

\subsection{Code architecture}
\label{subsec:code}

Describe how the code is written and how it works, the astropy.distribution thingy, MCMC for error determination etc.

% ------------------------------------------------------------------------------
\section{Analysis} 
\label{sec:analysis}

% ... explore a number of methods to infer properties of the bursting source,  
% based on the extent of available observational data, and assess their accuracy.

% seems to be a newer/alternative version of sec:results?

Here we describe different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.
%
We focus on adopting a suitable approach given the availability of data, which will differ for different sources. 
%
We also describe how the {\sc concord} code can be used to fully account for the astrophysical uncertainties, and as a companion to this section (and the code itself) we provide a companion {\sc jupyter} notebook [{\bf ref}] which replicates the analyses described below.

@@

While it may seem strange to base an analysis on the {\it non-}detection of bursts, practically even for the best-studied sources the X-ray observation duty cycle is of order a few percent \cite[e.g.][]{minbar}. There is a high probability that bursts will be missed, particularly if the accretion rate is low (and hence the bursts are infrequent). In extreme cases this can mean that no bursts whatsoever are observed,  which may also be explained by the compact object being a black hole rather than a neutron star.

@@

For example, based on {\it RXTE}/PCA observations of the accretion-powered millisecond pulsar IGR~J00291+5934, \cite{gal06b} derived joint constraints over the distance $d$ and fuel H-fraction $X_0$, adopting  the predictions of a simple numerical model ({\it ref to {\sc settle}}). The method is relatively simple\footnote{Although the method is straighforward, we choose not to implement it in {\sc concord}, as this is a relatively niche case. We include it as an example here for completeness.}; 
%
the persistent flux history is converted to an accretion rate history given an assumed distance (via eq. ??), and a burst train is generated. The times of the predicted bursts are compared to the good time intervals of the observations, and the simulation is repeated over a range of distances and times for the first burst. The likelihood for a given distance is then estimated as the fraction of simulated burst trains for which all the bursts fall within data gaps, and hence are missed.

Larger distance (or H-fraction) will imply larger burst rates, and hence a smaller likelihood of missing any. The constraints obtained suggest the distance is $\lesssim6$~kpc (at 3-sigma significance), provided $X_0\approx0.7$, a reasonable choice given the expectation of a H-rich donor in this $2.46$~hr binary orbital period system \cite[]{gal05a}.  
%
A separate argument is used to give a lower limit on the distance.

% moved the discussion to the results section
% ... and later moved it back --- dkg 9/21

%\subsubsection{The distance to IGR J00291+5934}
%
%Previously we used the absence of thermonuclear bursts in the accretion-powered millisecond pulsar IGR~J00291+5934 to constrain the distance (\S\ref{subsec:zerobursts}).
%%
In a subsequent outburst in 2015, a single burst was detected by {\it Swift}/XRT \cite[]{kuin15}, offering the opportunity to verify the previously-determined distance range. The estimated peak 
% 0.3--10~keV flux in the range (3.0--$5.6)\times10^{-11}\ \epcs$. 
0.1--35~keV flux was $(18\pm4)\times10^{-8}\ \epcs$ \cite[]{defalco17}, with the burst exhibiting spectral variations indicative of PRE. By comparing the peak flux to the empirical Eddington luminosity (see \S\ref{subsec:1burst}), the corresponding distance 
\cite[including the inclination range of 22--32$^\circ$ suggested by][]{torres08},
% possible inclinations as described in sec \S\ref{subsec:data}) is $4.4_{-0.7}^{+0.8}$~kpc, which is mostly inconsistent with the earlier limit. Clearly, such limits must be viewed with some caution.
is $5.0_{-0.5}^{+0.7}$~kpc, which is fully consistent with the previously established limit. The effect of the inclination constraints is illustrated in Fig. \ref{fig:dist_igr00291}).

\begin{figure}[ht!]
\plotone{fig1.pdf}
\caption{Estimated distance for the accretion-powered X-ray pulsar IGR~J00291+5934, based on the peak flux of the sole thermonuclear burst observed by \citet[][solid histograms]{defalco17};
%
and based on the {\it non-}detection of bursts by \citet[][red dashed lines]{gal06b}.
%
Note the good agreement between the methods.
%
The distribution for an isotropic distribution of system inclinations (up to a maximum value of $75^\circ$, motivated by the lack of dips in the X-ray intensity) is shown as the blue shaded histogram. Imposing the inclination constraint of 22--32$^\circ$ suggested by \cite{torres08}, for which the predicted burst anisotropy factor is $\xi_b=0.704$ on average, gives instead the orange-shaded histogram. The resulting $1\sigma$ confidence interval (green lines) is $5.0_{-0.5}^{+0.7}$~kpc.
\label{fig:dist_igr00291}}
\end{figure}

\subsection{One burst -- estimating the distance and burst rate} 
\label{subsec:1burst}

@@

For a single burst detected from a bursting source, provided the data quality and signal-to-noise are sufficient, we can expect to measure the burst peak flux, fluence, and timescale. If the burst exhibits PRE, we can infer the distance (\S\ref{subsec:distance}); from the persistent flux, the accretion rate; and (via burst models) the expected burst rate. The detailed shape of the lightcurve can provide constraints from comparisons with numerical models (see \S\ref{subsec:lccompare}), but here we focus on the more straightforward calculation which may be done from the simple burst measurements alone.

[{\bf detailed description of approach for distance here}]

The distance can be estimated from the peak flux of a PRE burst based on equation \ref{disteq}. The challenge is taking into account the uncertainties in the various parameters that go into this calculation. If we adopt the empirical Eddington luminosity of \cite[]{kuul03a}, the only remaining factor is the burst emission anisotropy $\xi_b$.
%
As for the case of the burst in IGR~J00291+5934, the approach adopted in {\sc concord} to take this into account is to generate a distribution of $\xi_b$ values, based on an isotropic distribution of inclinations within known limits. The presence (absence) of X-ray dips occurring at the binary orbital period will constrain the inclination to greater (less) than $75^\circ$ [{\bf ref}]. However, more stringent constraints may be available.

% [{\bf what if you don't want to use the empirical value? }]
The distance method in {\sc concord} can instead estimate the distance based on the theoretical Eddington luminosity (equation \ref{ledd}). In that case, values for $M_{\rm NS}$, $T_e$, $X$ and $1+z$ are assumed; these values, or distributions thereof, can also be provided by the user. No intrinsic uncertainty is otherwise assigned to the theoretical value; in contrast, the uncertainty in the empirical value will contribute to the overall distance uncertainty.

% moved from later section
% \subsubsection{System properties of IGR~J17591$-$2342}

We use as an example 
% of analysis of a single burst (\S\ref{subsec:1burst}) 
the event detected with \igr/JEM-X from IGR~J17591$-$2342 \cite[]{kuiper20}. This burst was observed  when the estimated persistent bolometric flux was $(1.2\pm0.2)\times10^{-9}\ \epcs$, and exhibited a fluence of $E_b =(1.1\pm0.1)\times10^{-6}\ {\rm erg\,cm^{-2}}$. The peak flux was $(7.6\pm1.4)\times10^{-6}\ \epcs$.

\cite{kuiper20} estimated the distance to the source based on the empirical value of the Eddington luminosity, and incorporating the expected anisotropy for an inclination range of 
% kuiper20 uses slightly more restrictive 28-30
24--30$^\circ$ \cite[]{sanna18} as $7.7_{-0.6}^{+0.8}$~kpc. Here the uncertainty is dominated by the peak flux, since the anisotropy factor is effectively fixed for such a narrow range of inclination, at $\xi_b=0.70$.
%
The PDF for the distance, which is distinctly non-Gaussian, is plotted in Figure \ref{fig:dist_igr17591}. We also show the corresponding distribution resulting from adoption of the theoretical Eddington luminosity, according to equation \ref{ledd}.

@@

\begin{figure}[ht!]
\plotone{fig2.pdf}
\caption{Distance estimates for IGR J17591$-$2342, based on the single PRE burst detected with JEM-X \cite[]{kuiper20}. The blue histogram and green 1-$\sigma$ limits replicate the analysis in that paper, comparing the flux to the empirical Eddington limit of \cite[]{kuul03a}, and adopting the inclination limis of \cite[]{sanna18}.  For comparison we also include the inferred distance distribution (orange histogram) adopting the theoretical Eddington luminosity, for a neutron-star mass of $1.4M_\odot$, and surface redshift of $1+z=1.259$.
\label{fig:dist_igr17591}}
\end{figure}

The {\sc concord} code allows us to pass the derived distance distribution directly to the method for calculating the accretion rate at the time of the burst, also incorporating the dependence of the persistent flux anisotropy (which is not the same as for the burst flux).
%
The persistent flux can then be converted to an accretion rate, given the estimated anisotropy factor predicted (separately) for the persistent emission, as
\begin{eqnarray}
\dot{m} & = L_{\rm pers} (1+z) (4\pi R^2(GM/R_{\rm NS}))^{-1}\\
& = \ldots
% TODO: give expanded expression for this and move to earlier section
\end{eqnarray}
via the {\tt mdot} method, and incorporating the estimated distance distribution already determined. 
%
Based on the distance constraints, the estimated accretion rate  is within the range $(2.0_{-0.4}^{+0.6})\times10^3\ \rm{g\,cm^{-2}\,s^{-1}}$ for the usual assumed range for neutron star mass and radius (and hence redshift).

@@

% here's the clever bit

Given a numerical burst ignition model, we could now go on to estimate the recurrence time, and check if it was consistent with the detection of only a single burst over the entire observed outburst. However, there is an extremely wide range of burst recurrence times at fixed accretion rate, depending upon the unknown fuel composition. we could potentially obtain constraints on the fuel composition by comparison of the observed lightcurve to simulations, but it's also possible that multiple different combinations of accretion rate and composition could give rise to similar composition at ignition (and hence shape of the lightcurve).
